Using Probabilistic Knockoffs of Binary Variables to Control the False Discovery Rate

### 2015-07-14 masters_paper/paper.tex
Draft of section 2

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

@@

\subsection{Binary Knockoffs}
    Though the ``original" knockoff method, described above, achieves the exact desired knockoff properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$. In particular, these values will often have very different empirical distributions. This effect is particularly noticeable when the random variable $x_j$ is discrete but $\tilde X_j$ does not even consist of the same set of values. The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics as linear regression which knockoffs are designed to hit, the original knockoffs will fail to control FDR. \par
    Thus, this paper offers a new method of generating knockoffs which should offer superior performance with other GLMs. In this new method, the matrix of knockoff features $\tilde X$ will be generated by row from a random vector valued variable $\mathbf{\tilde x}$, each entry $\tilde x_j$ of which will have the property $\tilde x_j \sim x_j$. This random variable will also abide by a relaxation of the original knockoff condition:
    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\hat\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
    This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
    Generating such knockoffs for a completely general case of any $\mathbf x$ is likely infeasible. However, in the specific case where $\mathbf x$ is a random binary vector variable, this paper will demonstrate a method to generate such knockoff variables. Since binary data is extremely common in many areas, this will hopefully serve as a useful tool.

@@

\section{Issues With Deterministic Knockoffs} 
    {\em Note:} I will try to hold to the convention that $X$ is the $n\times p$ data matrix, while $\bf x$ is the random vector variable from which each row of $X$ was drawn. Accordingly, $\tilde X$ will be the knockoff matrix while $\bf \tilde x$ is a random vector variable, at least when $\tilde X$ is generated randomly. $x_i$ will be the random variable corresponding to the $i$th entry of $\bf x$, while $X_i$ is $i$th column of $X$, with observations drawn from $x_i$. $\hat \Sigma = \frac{1}{n}X^TX$ is the empirical covariance matrix associated with $X$, while $\Sigma=\E(\mathbf x \mathbf x^T)$ is the theoretical covariance matrix associate with $\mathbf x$. Also, for a square matrix $A$, $\diag A$ is the vector of values along the diagonal, while for a vector $\mathbf a$, $\diag \mathbf a$ is a square diagonal matrix with $\mathbf a$ along the diagonal.\\

    Simulations with deterministic knockoffs reveal that they don't perform well in L1 regularized logistic regression. Even when $X_i$ is a null predictor of $y$, the $X_i$ still tend to enter the model prior to $\tilde X_i$. The issue is that even when $x\sim N_p({\bf 0}, \Sigma)$ for some $\Sigma\succeq 0$, the $\tilde X_i$ do not fit a normal distribution. This can be seen below in figure 1, where normal Q-Q plots of simulated variables $X_i$ and corresponding knockoffs $\tilde X_i$ are provided. Of course, when $X$ is a binary vector, $\tilde X$ completely doesn't match its distribution.
    \begin{figure}[h!]
        \includegraphics[width=20cm]{images/normalQQ}
        \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
    \end{figure}
    These

>>>>

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins, caption} 
\captionsetup{font={small,singlespacing},margin=1cm}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

@@

\subsection{Binary Knockoffs}
    Though the ``original" knockoff method, described above, achieves the exact desired knockoff properties, the individual values in the vector $X_j$ will have little relation to the individual values in $X_j$. In particular, these values will often have very different empirical distributions. This effect is particularly noticeable when the random variable $x_j$ is discrete but $\tilde X_j$ does not even consist of the same set of values. The end result is that for variable selection for other generalized linear models, which do not have the same sufficient statistics as linear regression which knockoffs are designed to hit, the original knockoffs will fail to control FDR. \par
    Thus, this paper offers a new method of generating knockoffs which should offer superior performance with other GLMs. In this new method, the matrix of knockoff features $\tilde X$ will be generated by row from a random vector valued variable $\mathbf{\tilde x}$, each entry $\tilde x_j$ of which will have the property $\tilde x_j \sim x_j$. This random variable will also abide by a relaxation of the original knockoff condition:
    \[ \E[\mathbf{\tilde x}^T\mathbf{\tilde x}]=\Sigma \; \textrm{ and } \; \E[\mathbf{\tilde x}^T \mathbf x \st \mathbf x] = \Sigma - \diag{\mathbf s}\]
    This will ensure that, at the very least, these knockoffs are a suitable replacement for the original knockoffs for linear regression. \par
    Generating such knockoffs for a completely general case of any $\mathbf x$ is likely infeasible. However, in the specific case where $\mathbf x$ is a random binary vector variable, this paper will demonstrate a method to generate such knockoff variables. Since binary data is extremely common in many areas, this will hopefully serve as a useful tool.

@@

\section{Issues With Deterministic Knockoffs} 
    The knockoff filter and $W$ statistics have a very natural extension to generalized linear models. Here, where $l(\beta\st X, \tilde X)$ is the log likelihood of coefficients $\beta$ given the model, the $L1$ regularized regression model will have estimated coefficient vector
    \[ \beta(\lambda) = \textrm{argmin}_\mathbf b \left\{l(\beta\st X, \tilde X) + \lambda\|b\|_1 \right\}\]
    The $Z_j$ and $\tilde Z_j$ are then once again the largest $\lambda$ such that the given original or knockoff feature has a positive coefficient, and then $W_j$ is constructed exactly as before. The problem with this is that if the given GLM doesn't have the same sufficient statistics as linear regression, there is no theoretical guarantee that the knockoff filter will control FDR. This is since, for null $X_j$, there is no theory to guarantee the $Z_j$ and $\tilde Z_j$ are exchangeable, and thus $W_j$ is not guaranteed to be symmetrically distributed around $0$. \par
    \begin{figure}[h]
        \begin{center}
        \includegraphics[width=15cm]{images/entryrate_original_logit}
    \end{center}
        \caption{The mean cumulative portion of $W$ statistic which are positive by absolute rank of the value. The mean is over 1000 simulations. On the left side, the features are drawn from a random standard normal distribution. On the right side, they are drawn as Bernoulli with probability $.25$ of being 1. The top has 50 features uncorrelated with the outcome, while the bottom has 40 features uncorrelated with the outcome and 10 correlated with it.}
    \end{figure}

    \FloatBarrier
    To test how bad this problem is, I simulated the method with the original knockoffs with logistic regression. The results can be seen in figure 1, which shows cumulatively, starting with the largest absolute value, what portion of the $W_j$ are positive. For the method to work, for the null predictors, this should be a consistent $50\%$, indicative of the desired symmetric distribution. This is particularly important for high values, which play the largest roll in selecting a threshold. We see that in the top left corner, when all features are null and normally distributed, that we almost achieve this, with only a small, but consistent, deviation from $50\%$. However, in the top right, when the predictors are binary, the null originals enter first far more often, which will the FDR too high. \par
    This gets much worse when some true predictors are included, in the bottom row of the figure. For both normal and binary predictors, the originals come in first far more often in the beginning. This will completely ruin FDR control, since the portion of null predictors will be assumed far lower than it actually is. It is worth noting that the particular shape of the curve in the diagram, with the plateau for the null predictors through 10, is partially an artifact of how the totals were tabulated. The mean through the first ten is the mean for null predictors that came in first out of both null and true predictors; since the largest ten $W_j$ will tend to be true predictors, the average over the null predictors represent the few extreme $W_j$ which managed to beat out some number of true predictors.  \par
    \begin{figure}[h]
        \includegraphics[width=15cm]{images/normalQQ}
        \caption{Normal Q-Q of original variables and knockoffs for simulation with $4$ variables and 1,000 observations}
    \end{figure}

    Upon examination, the issue seems to be that the original style knockoff features have a very different empirical distribution than the features they are imitating. This is obviously true with the binary features, but even with the normal features, the distributions differ a lot. In figure 2, the distribution of four feature and the original style knockoff are plotted against a normal distribution in Q-Q plots. As you can see, they don't match at all. This is suggestive of a solution; if the knockoff variables match the marginal distribution of the original features better, the result may be more suitable for FDR control in GLMs. \par
    \FloatBarrier
